---
title: "Regression Models Course Project"
output: html_document
---
Loading required packages

```{r results='hide', message=FALSE, warning=FALSE}
library("ggplot2")
library(dplyr)
library(Hmisc)
library(reshape2)
library(gridExtra)
```

Loading mtcars dataset and coverting the required variables to factors

```{r results='hide', message=FALSE, warning=FALSE}
data(mtcars)
mtcars <- mutate(mtcars,am = as.factor(am),cyl = as.factor(cyl),gear = as.factor(gear),
                 carb = as.factor(carb),vs = as.factor(vs))
```

Lets check if mpg is different for automatic and manual transimission cars using a T-test
Null hypothesis will be no difference in mean MPG between the two groups
p-value is less than 0.05,null hypothesis is rejected and we can confirm that mean mpg for automatic and manual cars are different
Estimated mean MPG of manual transmission cars is 7 more than automatic transmission cars 

```{r}
ttest_am <- t.test(mpg ~ am,data=mtcars)
ttest_am$p.value
ttest_am$estimate
```

To quantify the MPG difference we need to,
- explore the data to understand if there are other variables responsible for variation in mpg
- build regression models to estimate MPG difference based on am

Data exploration
- Box plot reveals that manual transmission cars have high MPG
- Automatic transmission cars tend to have high mean weight, high mean hp, low mean gear, high number of cylinders
- There is relationship b/w mpg and disp,hp,wt,drat
- Variation in MPG can not be explained only by transmission type
- correlation matrix reveals 

```{r warning=FALSE}
am_summary <- mtcars %>%
  group_by(am) %>%
  summarise(count = n(),cyl=mean(cyl)*100,disp=mean(disp),hp=mean(hp),drat=mean(drat)*100,wt=mean(wt)*100,
            qsec=mean(qsec)*10,vs=mean(vs)*1000,gear=mean(gear)*100,carb=mean(carb)*100) # scaling by a factor of 10/100 or 1000 to have all variables in the same range
t <- as.data.frame(t(am_summary))
t <- t[-1, ]
colnames(t) <- c("Automatic","Manual")
t$name <- rownames(t)
mdata <- melt(t,id=c("name"))
mdata$value1 <- as.numeric(as.character(mdata$value))

boxplot(mtcars$mpg ~ mtcars$am, xlab="Transmission (0 = Automatic, 1 = Manual)", ylab="MPG",
        main="Boxplot of MPG vs. Transmission")
ggplot(mdata,aes(name,value1,fill=variable))+
  geom_bar(position="dodge",stat="identity") 
plot1 <- qplot(wt,mpg,data = mtcars,color = am)
plot2 <- qplot(hp,mpg,data = mtcars,color = am)
plot3 <- qplot(drat,mpg,data = mtcars,color = am)
plot4 <- qplot(cyl,mpg,data = mtcars,color = am)
plot5 <- qplot(disp,mpg,data = mtcars,color = am)
grid.arrange(plot1,plot2,plot3,plot4,plot5,ncol=2)
```
```{r}
mtcars_corr_matrix <- rcorr(as.matrix(mtcars))
corr_mpg <- as.data.frame(mtcars_corr_matrix$r)
corr_mpg$name <- rownames(corr_mpg)
corr_mpg[corr_mpg$name == "mpg",-12]
```

Regression Models

- model based on am explains only 33.8% of variantion so more varibles need to be included to quantify mpg variation
```{r}
fit_am <- lm(mpg ~ am,data=mtcars)
summary(fit_am)$adj.r.squared
```
- model based on all varibles explains 78% of variation but none of the variables are statistically significant in explaining the relationship

```{r}
fit_allvars <- lm(mpg ~ .,data=mtcars)
summary(fit_allvars)$adj.r.squared
```
- using stepwise backward selection we end up on mpg ~ wt + qsec + am and this model explains 83% of variance
```{r results = 'hide'}
stepModel <- step(fit_allvars, k=log(nrow(mtcars)))
```
```{r}
stepModel$call
summary(stepModel)$adj.r.squared
```
- But there is an interaction between am and hp/wt/disp so building a model with interaction b/w wt & am
```{r}
fit_am_wt_interaction <- lm(mpg ~ wt + qsec + am + wt:am ,data=mtcars )
summary(fit_am_wt_interaction)$adj.r.squared
```
- Model selection
```{r results = 'hide'}
anova(fit_am, fit_allvars, stepModel, fit_am_wt_interaction)
confint(fit_am_wt_interaction) # results hidde
```
- Model with higher explainaibilty based on r-square is selected
```{r}
summary(fit_am_wt_interaction)$coef
```

Thus, the result shows that when “wt” (weight lb/1000) and “qsec” (1/4 mile time) remain constant, cars
with manual transmission add 14.079 + (-4.141)*wt more MPG (miles per gallon) on average than cars with
automatic transmission. That is, a manual transmitted car that weighs 2000 lbs have 5.797 more MPG than
an automatic transmitted car that has both the same weight and 1/4 mile time.

Residual Analysis and Diagnostics
Please refer to the Appendix: Figures section for the plots. According to the residual plots, we can verify
the following underlying assumptions:
- The Residuals vs. Fitted plot shows no consistent pattern, supporting the accuracy of the independence
assumption.
- The Normal Q-Q plot indicates that the residuals are normally distributed because the points lie closely
to the line.
- The Scale-Location plot confirms the constant variance assumption, as the points are randomly distributed.
- The Residuals vs. Leverage argues that no outliers are present, as all values fall well within the 0.5 bands
- As for the Dfbetas, the measure of how much an observation has effected the estimate of a regression coefficient,
we get the following result:

```{r}
sum((abs(dfbetas(fit_am_wt_interaction)))>1)
```

- Residual Plots

```{r}
par(mfrow = c(2, 2))
plot(fit_am_wt_interaction)
```
